\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2012-07-07 doc/USER_MANUAL/manual_SPECFEM3D.tex
updated the copyright information and a few things in the manual, to prepare for the release of version 2.1

\author{$\copyright$ Princeton University (USA) and\\
University of Pau / CNRS / INRIA (France)\\
Version 2.1 \\
"Sesame"}

@@

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Hom Nath Gharti,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Dennis McRitchie,
Matthias Meschede,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Zhinan Xie,
Hejun Zhu.\\
%\toall{add full list of developers here and update the AUTHORS file accordingly}

@@

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11}; it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\

@@

For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08,FiIgBuKe09,ViOp09}.\\

@@

The next release of the code will include support for GPU graphics card acceleration \citep{KoMiEr09,KoErGoMi10,MiKo10,Kom11}
as well as Convolutional or Auxiliary Differential Equation Perfectly Matched absorbing Layers (C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
It will also use the PT-SCOTCH parallel library for mesh partitioning.\\

SPECFEM3D includes coupled fluid-solid domains and
adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies so far.
To accommodate visco-plastic or non-linear rheologies, the reader can refer to 
the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.

@@

\item [Numerical simulations in general] ~\\
Forward simulations are described in detail in \cite{TrKoLi08,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}. Additional aspects dealing with adjoint simulations are described in
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10}.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.

@@

Princeton University, USA,
and University of Pau / CNRS / INRIA, France

$\copyright$ Princeton University and University of Pau / CNRS / INRIA, April 2011

@@

 MPI v 2.1 `Sesame' (Spectral ElementS on Any MEsh), July 2012: 
 Max Rietmann, Peter Messmer, Daniel Peter: support for CUDA GPUs. \\

 MPI v. 2.0 `Sesame' (Spectral ElementS on Any MEsh), November 2010:
 Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
 Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
 and Emanuele Casarotti, INGV Roma, Italy:
  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
  much faster solver using Michel Deville's inlined matrix products.\\

>>>>

\author{$\copyright$ Princeton University (USA) and\\
CNRS / INRIA / University of Pau (France)\\
Version 2.1}

@@

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Zhinan Xie,
Hejun Zhu.\\
%\toall{add full list of developers here and update the AUTHORS file accordingly}

@@

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin technique \citep{ReHi73,Arn82,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions \citep{WiStBuGh10,AcKo11}.
In particular, it can accurately handle very distorted mesh elements \citep{OlSe11}.\\

@@

For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11} and
in particular \citet{KoLiTrSuStSh04,LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07} and \citet{SeOl08}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation \citep{SaKoTr10}. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\

@@

This new release of the code (version 2.1) includes support for GPU graphics card acceleration \citep{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.\\

The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
It will also use the PT-SCOTCH parallel library for mesh partitioning.\\

SPECFEM3D includes coupled fluid-solid domains and
adjoint capabilities, which enables one to address seismological inverse problems, but for linear rheologies only so far.
To accommodate visco-plastic or non-linear rheologies, the reader can refer to
the \href{http://geoelse.stru.polimi.it/}{GeoELSE} software package \citep{CaGa97,StPaIg09,ChMoTsBaKrKaStKr10}.

@@

\item [Numerical simulations in general] ~\\
Forward and adjoint simulations are described in detail in \cite{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,KoGoErMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99} or \cite{KoVi98}.
Additional aspects dealing with adjoint simulations are described in
\cite{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores is shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11},

\item [GPU computing] ~\\
Computing on GPU graphics cards for acoustic or seismic wave propagation applications
is described in detail in \cite{KoMiEr09,KoGoErMi10,KoErGoMi10,MiKo10,Kom11}.

@@

Princeton University, USA, and CNRS / INRIA / University of Pau, France

$\copyright$ Princeton University and CNRS / INRIA / University of Pau, July 2012

@@

 MPI v 2.1, July 2012:
 Max Rietmann, Peter Messmer, Daniel Peter, Dimitri Komatitsch, Joseph Charles, Zhinan Xie: support for CUDA GPUs,
better CFL stability for the Stacey absorbing conditions. \\

 MPI v. 2.0, November 2010:
 Dimitri Komatitsch, Nicolas Le Goff, Roland Martin and Pieyre Le Loher, University of Pau, France,
 Jeroen Tromp and the Princeton group of developers, Princeton University, USA,
 and Emanuele Casarotti, INGV Roma, Italy:
  support for CUBIT meshes decomposed by SCOTCH, METIS or ZOLTAN;
  much faster solver using Michel Deville's inlined matrix products.\\
